Universals of word order result from optimization of grammars for efficient communication

### 2019-04-18 second-draft/3-paper-names.tex
figure

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics. % \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic

@@

%TODO at some point address head direction idea
%- Vennemann 1972, 1974ab, 1976

%-Hawkins 1980 limitations of head direction, also Hawkins 1982

@@

%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
Human languages vary in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  the object `letter', and an adpositional phrase translating to `to a friend'.
%In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
Yet, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
The verb stands at the beginning in Arabic, and at the end in Japanese.
Arabic expresses `to' by a \emph{preposition} (\emph{preceding} the noun `friend'); Japanese uses a \emph{postposition} (\emph{following} it). % and a postposition in Japanese, following `friend'.
%In Arabic, the verb stands at the beginning, followed by the objecat and the adpositional phrase.
%In Arabic, `to' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
%In Japanese, it is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
%It turns out that these are deep and stable regularities:

This variation reflects a deep and stable regularity:
While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
this is strongly correlated with the occurrence of pre- or postpositions (Figure~\ref{fig:corr-table}, top):
Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
%However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
%As Figure~\ref{fig:corr-table} shows, this correlation has an extremely strong empirical basis:
%In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization. %, over 90 \% of languages satisfy this generalization .

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{figures/corr-full.png}
%
%\centering
%\begin{dependency}[theme = simple]
%   \begin{deptext}[column sep=1em]
%%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
%	   katabt \& risāla \& li \& sadīq  \\
%%	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
%	   wrote \& letter   \& to \& friend  \\
%   \end{deptext}
%\end{dependency}
%\begin{dependency}[theme = simple]
%   \begin{deptext}[column sep=1em]
%      tomodachi \& ni \& tegami-o \& kaita \\
% %\textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
%	   friend \& to \& letter  \& wrote \\
%   \end{deptext}
%\end{dependency}
%
%    \centering
%    \includegraphics[width=0.4\textwidth]{figures/corr-table.png}
%
%%\begin{dependency}[theme=simple]
%%    \begin{deptext}
%%    I \& wrote \& a \& letter \& to \& a \& friend \\
%%    \end{deptext}
%%    \depedge{2}{1}{nsubj}
%%    \depedge{2}{4}{obj}
%%    \depedge{4}{3}{ }
%%    \depedge{4}{3}{ }
%%    \depedge{2}{5}{}
%%    \depedge{5}{7}{ }
%%    \depedge{7}{6}{ }
%%\end{dependency}
	\caption{ The concept of a word order correlation: Languages can order the object after (Arabic) or before (Japanese) the verb, and have prepositions (Arabic) or postpositions (Japanese). For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structure~\cite{wals}. Combinations on the diagonal are vastly more common than off-diagonal ones. % A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.
	}\label{fig:arabic-japanese-simple}	\label{fig:corr-table}
\end{figure}

This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
These describe correlations between the relative positions of different types of expressions across languages.
The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of the adposition (`to'). %adpositional phrases (`to a friend') relative to the verb.
Greenberg also found that the order between verb and object is correlated with other aspects of a language's word order (Table~\ref{table:corr-dryer}), such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).
%Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
%The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.

@@

%TODO can we find citations from all camps?

@@

%Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing. These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.

% alternatively: expressivity vs simplicity

Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between the Forces of \emph{Unification} and \emph{Diversification}.
The Force of Unification is a pressure to reduce the complexity of the language to make production and processing as easy as possible.
Conversely, the Force of Diversification favors languages that provide different utterances for different meanings, so that the listener can unambiguously identify the meaning from the utterance.
%Minimizing speaker and listener effort represent two opposing forces:
%A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
%A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
These two forces act in opposing directions:
Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required in order to provide higher informativity.
The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in theoretical studies of language~\cite{gabelentz1901sprachwissenschaft,horn1984toward, lindblom1990explaining, schwartz1997dispersion, hawkins2004efficiency}.

Recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language, showing that they predict
both basic statistical properties of languages~\cite{ferreri2003least} and language development~\cite{smith2013linguistic},
and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.

While it has been suggested that these ideas should also apply to grammar~\cite{hawkins2004efficiency}, testing these accounts is more difficult in the domain of grammar, as this requires both large amounts of data representative of language use across languages and cultures, and computational methods for estimating the efficiency of entire languages.
%TODO somewhere start calling this Efficiency
%TODO explain why surprisal. some production/speaker work. (plan reuse)
%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
%This theory blabla Horn RSA
%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
In this work, we address these challenges by combining large-scale text data from 51 languages with state-of-the-art machine learning techiques to estimate the communicative efficiency of grammar.
We apply this approach to the problem of explaining Greenberg's word order correlation universals.
We first show that the word order of languages has evolved to optimize efficiency.
We then show that efficiency optimization accounts for the Greenbergian word order correlations, and makes fine-grained per-language predictions about individual syntactic patterns.
Our results show how general properties of efficient communication gives rise to the universal word order properties of human language.
%We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.

%In this work, we first provide a formalization of efficiency in the domain of word order. %, drawing on information-theoretic notions.

%In this work, we show computationally, using corpus data from 51 languages, that efficiency optimization accounts for the Greenbergian word order correlations.
%That is, we show that the word orders in natural languages have evolved to optimize efficiency, and that this optimization accounts for the prevalence of the word order correlations. %word orders in natural language

%OR say DLM and efficiency, as in the talks

%\section{DLM}

%
%\begin{figure}
%    \centering
%    \includegraphics[width=0.45\textwidth]{figures/letter-options.png}
%	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure. In this example, Grammar 1 achieves the lowest overall dependency length. Indeed, this is the }
%	\label{fig:pareto-plane}
%\end{figure}
%
%
%\begin{tabular}{|ll|c|cccc|llllllll}
%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Eff}  \\ 
%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%	& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}       }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}   }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}     }  \\
%	\emph{to}            & \emph{a friend} &&&&&\\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}       }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}   }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}   }  \\
%	\emph{is}        & \emph{a friend}  &&&&&\\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}       }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}     }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  }  \\
%	\emph{has}          & \emph{written}  &&&&&\\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  }  \\
%	\emph{friend} &  \emph{of John}  &&&&&\\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf}  }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  }  \\
%	\emph{books} & \emph{that you read}  &&&&&\\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}     }
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf} }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  }  \\
%	\emph{that} & \emph{she has arrived}  &&&&&\\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}     }
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  }  \\
%	\emph{went} & \emph{to school}  &&&&&\\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}     }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}  }  \\
%	\emph{wants}   &  \emph{to leave}  &&&&&\\ \hline
%    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
%For each correlation, we provide an English example.
%\begin{table*}
%\begin{tabular}{|ll|l|l|l|ll|l|llllllll}
%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}   $^{***}$    }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}  $^{***}$ }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf} $^{***}$    } \\
%	\emph{to}            & \emph{a friend} \\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}   $^{***}$    }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}  $^{***}$ }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf} $^{**}$  }  \\
%	\emph{is}        & \emph{a friend} \\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}   $^{***}$    }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}  $^{***}$   }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}$^{**}$  }  \\
%	\emph{has}          & \emph{written} \\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} $^{***}$    }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} $^{***}$ }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}$^{***}$  }  \\
%	\emph{friend} &  \emph{of John} \\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf} $^{***}$    }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf} $^{**}$ }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}$^{***}$  }  \\
%	\emph{books} & \emph{that you read} \\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf} $^{***}$    }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}  $^{**}$   }
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf}$^{***}$ }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}$^{**}$  }  \\
%	\emph{that} & \emph{Mary has arrived} \\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf} $^{***}$    }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}  $^{***}$   }
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}$^{**}$  }  \\
%	\emph{went} & \emph{to school} \\ \hline
%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} $^{***}$    }
%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf} $^{**}$    }
%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} $^{***}$ }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}$^{***}$  }  \\

>>>>

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics. % \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
% TODO mention universals here
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic 
Specifically, we show that efficiency optimization explains  linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
% TODO change wording so it's clear that this is a very major thing

% Things TODO
% - make clear that efficiency is something both functional and generative linguists have been appealing to
% - at some point address head direction idea
%TODO at some point address head direction idea
%    - Vennemann 1972, 1974ab, 1976
%    -Hawkins 1980 limitations of head direction, also Hawkins 1982

@@

@@

%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
Human languages vary in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  the object `letter', and an adpositional phrase translating to `to a friend'.
%In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
Yet, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
The verb stands at the beginning in Arabic, and at the end in Japanese.
Arabic expresses `to' by a \emph{preposition} (\emph{preceding} the noun `friend'); Japanese uses a \emph{postposition} (\emph{following} it). % and a postposition in Japanese, following `friend'.

This variation reflects a deep and stable regularity:
While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
this is strongly correlated with the occurrence of pre- or postpositions (Figure~\ref{fig:corr-table}, top):
Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{figures/corr-full.png}
\caption{ The concept of a word order correlation: Languages can order the object after (Arabic) or before (Japanese) the verb, and have prepositions (Arabic) or postpositions (Japanese). For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structure~\cite{wals}. Combinations on the diagonal are vastly more common than off-diagonal ones. % A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.
	}\label{fig:arabic-japanese-simple}	\label{fig:corr-table}
\end{figure}

This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
These describe correlations between the relative positions of different types of expressions across languages.
The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of the adposition (`to'). %adpositional phrases (`to a friend') relative to the verb.
Greenberg also found that the order between verb and object is correlated with other aspects of a language's word order (Table~\ref{table:corr-dryer}), such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).

@@

@@

Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between the Forces of \emph{Unification} and \emph{Diversification}.
The Force of Unification is a pressure to reduce the complexity of the language to make production and processing as easy as possible.
Conversely, the Force of Diversification favors languages that provide different utterances for different meanings, so that the listener can unambiguously identify the meaning from the utterance.
These two forces act in opposing directions:
Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required in order to provide higher informativity.
The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in theoretical studies of language~\cite{gabelentz1901sprachwissenschaft,horn1984toward, lindblom1990explaining, schwartz1997dispersion, hawkins2004efficiency}.

Recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language, showing that they predict
both basic statistical properties of languages~\cite{ferreri2003least} and language development~\cite{smith2013linguistic},
and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.

While it has been suggested that these ideas should also apply to grammar~\cite{hawkins2004efficiency}, testing these accounts is more difficult in the domain of grammar, as this requires both large amounts of data representative of language use across languages and cultures, and computational methods for estimating the efficiency of entire languages.
In this work, we address these challenges by combining large-scale text data from 51 languages with state-of-the-art machine learning techiques to estimate the communicative efficiency of grammar.
We apply this approach to the problem of explaining Greenberg's word order correlation universals.
We first show that the word order of languages has evolved to optimize efficiency.
We then show that efficiency optimization accounts for the Greenbergian word order correlations, and makes fine-grained per-language predictions about individual syntactic patterns.
Our results show how general properties of efficient communication gives rise to the universal word order properties of human language.

\begin{table}
	\begin{center}
%		\setlength\tabcolsep{1.5pt}
\begin{tabular}{|c|ll|ll|}
	\hline
	%&
	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Hindi, ...)} \\ \hline\hline
	& \multicolumn{2}{c}{Correlates with...} & \multicolumn{2}{|c|}{Correlates with...}  \\
	&	Verb & Object     & Object & Verb    \\ \hline\hline
	& kataba & rasaa'il 	&tegami-o & kaita \\
	&	\emph{wrote} & \emph{letters} & 	\emph{letter} & \emph{wrote} \\ \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    sadiiq       &	tomodachi & ni \\
	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&kaana    &    sadiiq         &	tomodachi & datta \\
	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak & udesho \\
	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& sadiiq  &    John    & 	John no & tomodachi \\
	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uha       & 	anata-ga yonda & hon \\
	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tusil        & 	toochaku suru & koto \\
	&	\emph{that} & \emph{she arrives}  &	\emph{has.arrived} & \emph{that} \\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'ila lmadrasa     & 	gakkoo ni & ittekimashita \\
	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'uriid    &    'an 'ughaadir         & 	yom & itai \\
	& \emph{wants}   &  \emph{to leave}  &	\emph{to leave} & \emph{want} \\ \hline \hline
\end{tabular}
	\end{center}
	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples: Across the world, the orders of different constituents are strikingly correlated with that of verb and object. See Materials and Methods.}\label{table:corr-dryer}
\end{table}
